---
title: "task3"
author: "Mattia Storero"
date: "24/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
rm(list = ls())
```

```{r, include=FALSE}
library(tidyverse)
library(gapminder)
library(readr)
library(data.table)
library(plyr)
library(kableExtra)
```
## Background
The data were taken from "https://www.cdc.gov/nchs/data_access/vitalstatsonline.htm". The dataframes were downloaded, setted and selected thanks to the two guides.

```{r, include=FALSE}
#set the databases and selecting rows and columns looking at the guides
setwd("C:\\Users\\store\\Downloads\\Nat2018us")

data2018 <-
  read_fwf("Nat2018PublicUS.c20190509.r20190717.txt",
           fwf_empty("Nat2018PublicUS.c20190509.r20190717.txt")
           )
data2018 <-
  read_fwf("Nat2018PublicUS.c20190509.r20190717.txt",
           fwf_positions(start = c(475,504,13),
                         end = c(475,507,14),
                         col_names = c("sex","weight", "month")
                         )
           )

setwd("C:\\Users\\store\\Downloads\\Nat2017us")

data2017 <-
  read_fwf("Nat2017PublicUS.c20180516.r20180808.txt",
           fwf_positions(start = c(475,504,13),
                         end = c(475,507,14),
                         col_names = c("sex","weight", "month")
                         )
           )


```

## Working with the data
The data were cleaned fixing the variables and adding the column year for each database. After that, the two databases were merged together vertically in order to work with them.

```{r, include=FALSE}
#cleaning and fix the variables + adding the column year for each database

data2017$weight <- as.numeric(data2017$weight)
data2017$month <- as.numeric(data2017$month)
data2017$weight[data2017$weight == 9999] <- NA
data2017$sex <- as.factor(data2017$sex)
data2017$year <- 2017


data2018$weight <- as.numeric(data2018$weight)
data2018$month <- as.numeric(data2018$month)
data2018$weight[data2018$weight == 9999] <- NA
data2018$sex <- as.factor(data2018$sex)
data2018$year <- 2018


```


```{r, include=FALSE}
#merging vertically the two dataframes
total_data <- rbind(data2017, data2018)

```

## Analysis and answering to questions

1-The first point was to find out the proportion of boys and girls. For this reason, we worked with the variable "sex".
```{r, echo=FALSE, warning=FALSE}
#finding the proportions compared to the total and divide them
mean(data2017$sex == "M")
mean(data2017$sex == "F")
mean(data2017$sex == "M") / mean(data2017$sex == "F")

mean(data2018$sex == "M")
mean(data2018$sex == "F")
mean(data2018$sex == "M") / mean(data2018$sex == "F")

```

It's easy to see that the two values of boys/girls has changed, but not in a relevant way. In fact, it decreases from 1.047996 to 1.045748. To sum up, the number of girls has increased.

2-The second point was to find out the avarage weight in grams and to see if there were differences between boys and girls in 2018 thanks to a density plot.
```{r, include=FALSE}
#total average
average <- mean(total_data$weight, na.rm = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#density plot

ggplot(data2018, aes(x=weight, color=sex, fill=sex)) + 
 geom_histogram(aes(y=..density..), alpha=0.5, 
                position="identity")+
 geom_density(alpha=.2)+labs(title="Weight density curve",x="Weight(grams)", y = "Density")




```

3-The third point was to find out if the average birth weight vary by month and year.

```{r, include=FALSE}
mean(data2017$weight, na.rm = TRUE)
mean(data2018$weight, na.rm = TRUE)
weight_rel <- total_data %>%
  group_by(month,year) %>%
  summarise(average = mean(weight,na.rm = TRUE))
```


```{r, echo=FALSE, warning=FALSE}
as.data.table(data2017)[, .(avg_weight = mean(weight, na.rm = TRUE)), by = month]
as.data.table(data2018)[, .(avg_weight = mean(weight, na.rm = TRUE)), by = month]

```


It's possible to see that the annual average weight decreases from 3262.134 to 3261.649, whereas between the months there are relatively bigger differences (but not really relevant).





